Effect of quality-adjusted production function inputs

### 2015-06-02 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
New tables of Dmse and Dlambda.

<<setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE>>=
require(EconData)
require(EconModels)
require(dplyr)
require(reshape2)
require(ggplot2)
require(mosaic)

@@

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE     
)
@

@@

@@

lag.to.use <- 1 # years
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), 
                        f = function(model, id, lag=lag.to.use, pad="tail"){
print(lag)
  df <- dynamicSR(model, id=id, lag=lag, pad=pad)
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  df$Year <- model$data[["Year"]]
  return(df)  
}) %>% do.call(plyr::rbind.fill, .)

@@

We can test our hypothesis that quality-adjusting the factors of production
and including energy will reduce the Solow residuals~($\lambda$)
and the fitting residuals~($r$) by plotting
$\Delta \lambda_i$ vs. $\Delta r_i$, where
%
\begin{equation} \label{eq:delta_lambda}
  \Delta \lambda_i \equiv \left| \lambda_i \right| - \left| \lambda_{i, \, Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_r}
  \Delta r_i \equiv \left| r_i \right| - \left| r_{i, \,Unadjusted, \, Without \, energy} \right|.
\end{equation}
%
If most of the points fall in the lower-left quadrant of a graph, 
we can conclude that both Solow residual~($\lambda$) and
fitting residual~($r$) have decreased due to either 
quality-adjustment of the factors of production or 
including energy.

@@

We can also calculate the average values for $\Delta \lambda$ and $\Delta r$ 
as shown in Table~\ref{tab:traj.means.CD}.
%
<<TrajMeansTableFunction>>=
trajMeansTable <- function(data, caption, label){
  table <- xtable::xtable(data, caption=caption, label=label, digits=6)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE)
}
@

<<Trajectory_Means_CD, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CD") %>% select(-nest), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
               label="tab:traj.means.CD")
@
%
These results for the Cobb-Douglas model indicate that 
we should perform the same analysis for the CES model.

<<Solow_residual_trajectories_CESkle, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($kl$)$e$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="kle")))
@

<<Trajectory_Means_CESkle, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="kle")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($kl$)$e$ nesting.",
               label="tab:traj.means.CESkle")
@

@@

<<Trajectory_Means_CESlek, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="lek")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($le$)$k$ nesting.",
               label="tab:traj.means.CESlek")
@

@@

<<Trajectory_Means_CESekl, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="ekl")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($ek$)$l$ nesting.",
               label="tab:traj.means.CESekl")
@

>>>>

<<setup, echo=FALSE, message=FALSE, eval=TRUE, warning=FALSE>>=
require(EconData)
require(EconModels)
require(dplyr)
require(reshape2)
require(ggplot2)
require(mosaic)
require(xtable)

@@

opts_chunk$set(
  # ********************************************************
  eval=TRUE,
  # ********************************************************
  # tikz allows LaTeX code in graphical output. 
  # E.g., "$y$" for a variable name in the legend for a graph.  
  # dev='tikz',
  tidy=FALSE,
  comment=NA,
  # Tells whether to cache output from chunks are saved. Cacheing saves time. 
  # However, references to LaTeX tables DO NOT WORK if cacheing is turned on.
  cache=FALSE,
  warning=FALSE,  #Tells whether to show warnings in the output.
  # Turns off messages for all chunks. 
  # Set TRUE on an individual chunk to see it.
  message=FALSE, 
  # Tells whether to echo code for all chunks. 
  # Set TRUE on an individual chunk to see its code.
  echo=FALSE,
  # Sets all table captions to the top.
  xtable.caption.placement="top"
)
@

@@

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease the fitting residual and dynamic Solow residual?} 
\label{sec:quality_adj_energy_and_dynamic_solow_residual}
%++++++++++++++++++++++++++++++

We can test our hypothesis that quality-adjusting the factors of production
and including energy will both reduce the Solow residuals~($\lambda$)
and improve the fit to historical data
(thereby, reducing the fitting residuals, $r$)
by calculating 
$\Delta \lambda$ and $\Delta mse$, where
%
\begin{equation} \label{eq:delta_lambda}
  \Delta \lambda \equiv \left| \lambda \right| - \left| \lambda_{Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_mse}
  \Delta mse \equiv \frac{1}{N} \sum_i{r_i^2} - mse_{Unadjusted, \, Without \, energy}.
\end{equation}
%
where~$N$ is the number of years of data.
When $\Delta \lambda$ and $\Delta mse$ are negative, 
we observe reduction in the Solow residual ($\lambda$) and
the fitting residuals ($r$).

Table~\ref{tab:DSRmse_CD} shows the results for the Cobb-Douglas models.
%
<<MyXTableFunction>>=
myXTable <- function(data, caption, label, digits=2){
  table <- xtable(data, caption=caption, label=label, digits=digits)
  print(table, 
        caption.placement="top", 
        size="\\tiny", 
        include.rownames=FALSE, 
        math.style.negative=TRUE)
}
@
%
<<DSRmse_CD, results="asis">>=
myXTable(filter(AllCoeffs, model=="CD") %>% select(Country, model, flavor, energy, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for Cobb-Douglas models.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
Tables~\ref{tab:DSRmse_CESkle}--\ref{tab:CSRmse_CESekl} show results for the various CES nestings.
%
<<DSRmse_CESkle, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="kle")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($kl$)$e$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESlek, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="lek")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($le$)$k$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@
%
<<DSRmse_CESekk, results="asis">>=
myXTable(filter(AllCoeffs, model=="CES" & (nest=="kl" | nest=="ekl")) %>% 
           select(Country, model, flavor, energy, nest, Dlambda, Dmse),
         caption="$\\Delta \\lambda$ and $\\Delta mse$ for CES models with ($ek$)$l$ nesting.",
         label=paste0("tab:", opts_current$get("label")), 
         digits=6)
@

@@

lag.to.use <- 1 # years
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), 
                        f = function(model, id, lag=lag.to.use, pad="tail"){
  df <- dynamicSR(model, id=id, lag=lag, pad=pad)
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  df$Year <- model$data[["Year"]]
  return(df)  
}) %>% do.call(plyr::rbind.fill, .)

@@

We can test our hypothesis that quality-adjusting the factors of production
and including energy will reduce the Solow residuals~($\lambda$)
and the fitting residuals~($r$) by plotting
$\Delta \lambda_i$ vs. $\Delta r_i$, where
%
\begin{equation} \label{eq:delta_lambda_i}
  \Delta \lambda_i \equiv \left| \lambda_i \right| - \left| \lambda_{i, \, Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_r}
  \Delta r_i \equiv \left| r_i \right| - \left| r_{i, \,Unadjusted, \, Without \, energy} \right|.
\end{equation}
%
If most of the points fall in the lower-left quadrant of a graph, 
we can conclude that both Solow residual~($\lambda$) and
fitting residual~($r$) have decreased due to either 
quality-adjustment of the factors of production or 
including energy.

@@

We can also calculate the average values for $\Delta \lambda$ and $\Delta r$ 
as shown in Table~\ref{tab:traj.means.CD}.
%
<<Trajectory_Means_CD, results="asis">>=
myXTable(data = filter(traj.data.means, model=="CD") %>% select(-nest), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
         label="tab:traj.means.CD",
         digits=6)
@
%
These results for the Cobb-Douglas model indicate that 
we should perform the same analysis for the CES model.
%
<<Solow_residual_trajectories_CESkle, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($kl$)$e$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="kle")))
@
%
<<Trajectory_Means_CESkle, results="asis">>=
myXTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="kle")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($kl$)$e$ nesting.",
         label="tab:traj.means.CESkle",
         digits=6)
@

@@

<<Trajectory_Means_CESlek, results="asis">>=
trajMeansTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="lek")), 
               caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($le$)$k$ nesting.",
               label="tab:traj.means.CESlek",
               digits=6)
@

@@

<<Trajectory_Means_CESekl, results="asis">>=
myXTable(data = filter(traj.data.means, model=="CES" & (nest=="kl" | nest=="ekl")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($ek$)$l$ nesting.",
         label="tab:traj.means.CESekl",
         digits=6)
@
